---
title: 'Class 12: Introduction to Genome Informatics Lab'
author: 'Kevyn Aguilar Ramirez (PID: A16321291)'
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1. Proportion og G/G in population

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.)/nrow(mxl) * 100
```


Let's look at different population (GBR)
```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```


Find proportion of G|G

```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) * 100, 2)
```

This variant that is associated with childhood asthma is more frequent in the GBR population than the MXL population.

Let's now dig into this further

## Section 4: Population Scale Analysis
One sample is obviously not enough to know what is happening in a population. You are
interested in assessing genetic differences on a population scale. 

How many samples do we have?
> Q13: Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes. 

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
nrow(expr)
```

```{r}
table(expr$geno)
```
> The sample sixe for A/A is 108, 233 for A/G, and 121 for G/G.

```{r}
summary(expr)
```
> The median expression level is 25.116.

> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

```{r}
library(ggplot2)
```

Let's make a boxplot

```{r}
ggplot(expr) + aes(x=geno, y=exp, fill=geno) +
  geom_boxplot(notch = TRUE)
```
> The expression for G/G tends to be underexpressed/lower than expression for A/A. It is possible that the SNP affects the expression of ORMDL3.

